<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "README-"
)
options(width = 95)
```

utf8
====

[![Build Status (Linux)][travis-badge]][travis]
[![Build Status (Windows)][appveyor-badge]][appveyor]
[![Coverage Status][codecov-badge]][codecov]
[![CRAN Status][cran-badge]][cran]
[![License][apache-badge]][apache]
[![CRAN RStudio Mirror Downloads][cranlogs-badge]][cran]


*utf8* is an R package for manipulating and printing UTF-8 text that fixes
[multiple][windows-enc2utf8] [bugs][emoji-print] in R's UTF-8 handling.


Installation
------------

### Stable version

*utf8* is [available on CRAN][cran]. To install the latest released version,
run the following command in R:

```r
### install.packages("utf8") # not available yet, actually
```


### Development version

To install the latest development version, run the following:

```r
tmp <- tempfile()
system2("git", c("clone", "--recursive", shQuote("https://github.com/patperry/r-utf8.git"), shQuote(tmp)))
devtools::install(tmp)
```

Note that *utf8* uses a git submodule, so you cannot use
`devtools::install_github`.


Usage
-----

### Validate character data and convert to UTF-8

Use `as_utf8` to validate input text and convert to UTF-8 encoding. The
function fails if the input text has the wrong declared encoding:

```{r}
# second entry is encoded in latin-1, but declared as UTF-8, so
x <- c("fa\u00E7ile", "fa\xE7ile", "fa\xC3\xA7ile")
Encoding(x) <- c("UTF-8", "UTF-8", "bytes")
as_utf8(x) # fails
```

### Normalize data

Use `utf8_normalize` to convert to Unicode composed normal form (NFC).
Optionally apply compatibility maps for NFKC normal form or case-fold.

```{r}
# three ways to encode an angstrom character
(angstrom <- c("\u00c5", "\u0041\u030a", "\u212b"))
utf8_normalize(angstrom) == "\u00c5"

# perform full Unicode case-folding
utf8_normalize("Größe", map_case = TRUE)
```

### Print emoji

On some platforms (including MacOS), the R implementation of `print` uses an
outdated version of the Unicode standard to determine which characters are
printable. Use `utf8_print` for an updated print function:

```{r}
print(intToUtf8(0x1F600 + 0:79)) # with default R print function

utf8_print(intToUtf8(0x1F600 + 0:79)) # with utf8_print, truncates line

utf8_print(intToUtf8(0x1F600 + 0:79), chars = 1000) # higher character limit
```

Citation
--------

Cite *utf8* with the following BibTeX entry:

    @Manual{,
        title = {utf8: UTF-8 Text Processing},
        author = {Patrick O. Perry},
        year = {2017},
        note = {R package version 1.0.0},
    }


Contributing
------------

The project maintainer welcomes contributions in the form of feature requests,
bug reports, comments, unit tests, vignettes, or other code.  If you'd like to
contribute, either

 + fork the repository and submit a pull request (note the nonstandard
   instructions for [building from source][building]);

 + [file an issue][issues];

 + or contact the maintainer via e-mail.

This project is released with a [Contributor Code of Conduct][conduct],
and if you choose to contribute, you must adhere to its terms.


[apache]: https://www.apache.org/licenses/LICENSE-2.0.html "Apache License, Version 2.0"
[apache-badge]: https://img.shields.io/badge/License-Apache%202.0-blue.svg "Apache License, Version 2.0"
[appveyor]: https://ci.appveyor.com/project/patperry/r-utf8/branch/master "Continuous Integration (Windows)"
[appveyor-badge]: https://ci.appveyor.com/api/projects/status/github/patperry/r-utf8?branch=master&svg=true "Continuous Inegration (Windows)"
[codecov]: https://codecov.io/github/patperry/r-utf8?branch=master "Code Coverage"
[codecov-badge]: https://codecov.io/github/patperry/r-utf8/coverage.svg?branch=master "Code Coverage"
[conduct]: https://github.com/patperry/r-utf8/blob/master/CONDUCT.md "Contributor Code of Conduct"
[cran]: https://cran.r-project.org/package=utf8 "CRAN Page"
[cran-badge]: http://www.r-pkg.org/badges/version/utf8 "CRAN Page"
[cranlogs-badge]: http://cranlogs.r-pkg.org/badges/utf8 "CRAN Downloads"
[emoji-print]: https://twitter.com/ptrckprry/status/887732831161425920 "MacOS Emoji Printing"
[issues]: https://github.com/patperry/r-utf8/issues "Issues"
[travis]: https://travis-ci.org/patperry/r-utf8 "Continuous Integration (Linux)"
[travis-badge]: https://api.travis-ci.org/patperry/r-utf8.svg?branch=master "Continuous Integration (Linux)"
[windows-enc2utf8]: https://twitter.com/ptrckprry/status/901494853758054401 "Windows enc2utf8 Bug"
